---
title: "Initial Descriptives"
author: "Samuel Snelson"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(geofacet)
theme_set(theme_light(base_family = "Avenir"))
```




```{r}

d <- read.csv("final_report/processed_data/child_welfare_data.csv")

# setting path for figures to be saved
fig_path <- "final_report/figures/"

```


First we'll get a sense of just how many youth are in foster care across states and year. 

There are a few measures of the number of youth in foster care in a state and year. We'll check them against each other. 

We have `n_first_day_count` which is a measure of the number of youth in care on the first day of the fiscal year, October 1. Another is `n_in_care_check` based on the percentage of youth maltreated in foster care in a given year and state. Another is `n_in_care_place` which is based on the percentage of youth who experience different placement frequencies (this one has to be grouped by state and year). 

```{r}

# let's produce some figures showing the trends in foster care populations across these states

# setting up options for map plots
map_options <-  
  theme(axis.title.x = element_text(size = 18, 
                                    face = "bold"), 
        axis.title.y = element_text(size = 18, 
                                    face = "bold"), 
        plot.caption = element_text(size = 18),
        strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 14), 
        axis.text.x = element_text(size = 14, 
                                   angle = 45, 
                                   vjust = 0.45), 
        plot.title = element_text(size = 32, 
                                  vjust = 1, 
                                  face = "bold"))


p_fc_pop_overall <- d |>
  select(year, n_in_care_total) |> 
  distinct() |> 
  group_by(year) |> 
  summarize(total_pop = sum(n_in_care_total)) |> 
   
  ggplot(aes(year, total_pop)) +
  geom_line(linewidth = 1.25, 
            alpha = 0.5) + 
  geom_label(aes(label = format(total_pop, 
                                big.mark = ",")), 
             position = position_dodge(0.25), 
             color = "grey25") +
  
  labs(title = "Total Population of Foster Care in US, 2017-2021", 
       x = "Year", 
       y = "Number of Foster Youth", 
       caption = "Source: US Children's Bureau Child Welfare Outcomes Report Data") + 
  theme(plot.title = element_text(face = "bold"), 
        axis.title.x = element_text(face = "bold"), 
        axis.title.y = element_text(face = "bold"))


# Overall Foster Care Population 
ggsave(paste0(fig_path, "foster_care_population_overall.png"),
       plot = p_fc_pop_overall, 
       width = 7, height = 5, units = "in")

```

Now let's look at how the foster care population has changed over time


```{r}

p_fc_pop_states <- d |> 
  distinct(state, year, n_in_care_total) |> 
  filter(state != "Puerto Rico") |>
  mutate(n_in_care_scale = n_in_care_total / 1000) |> 
  ggplot(aes(year, n_in_care_scale)) + 
  geom_line(color = "firebrick") + 
  facet_geo(~ state, grid = "us_state_grid1", 
            scales = "free_y") + 
  
  labs(title = "Total Population of Foster Care in US, 2017-2021", 
       x = "Year", 
       y = "Number of Youth in Foster Care (1000s)", 
       caption = "Source: US Children's Bureau Child Welfare Outcomes Report Data") + 
  map_options


# Foster Care Population by State
ggsave(paste0(fig_path, "foster_care_population_bystate.png"), 
       plot = p_fc_pop_states, 
       width = 20, height = 12, units = "in")


# The scaling of the y-axis and size of text is a bit awkward, so we'll return to this bit this is the bulk of the plot we're interested in




```


Let's now turn to descriptive statistics on maltreatment in care and placement instability (maybe both aggregated and by length in care)


```{r}

# lollipop plot for maltreatment in care 

p_maltreat_in_care_lolli <- d |> 
  select(state, year, region, pct_maltreat_in_care) |> 
  distinct() |> 
  ggplot(aes(pct_maltreat_in_care, 
             reorder(state, pct_maltreat_in_care))) + 
  geom_linerange(aes(xmin = 0, 
                     xmax = pct_maltreat_in_care), 
                 color = "firebrick", 
                 linewidth = 2) +
  geom_point(size = 6) + 
  facet_grid(~ year) + 
  labs(title = "Prevalence of Maltreatment in Foster Care, 2017-2021", 
       x = "Percent Maltreated",
       y = "State", 
       caption = "Source: US Children's Bureau Child Welfare Outcomes Report Data") + 
  theme(plot.title = element_text(size = 32, 
                                  face = "bold"), 
        axis.title.x = element_text(size = 20, 
                                    face = "bold"), 
        axis.title.y = element_text(size = 20, 
                                    face = "bold"), 
        axis.text.x = element_text(size = 20), 
        axis.text.y = element_text(size = 20, 
                                   face = "bold"), 
        strip.text = element_text(size = 20), 
        plot.caption = element_text(size = 20))

ggsave(paste0(fig_path, "pct_maltreated_lollipop.png"), 
       plot = p_maltreat_in_care_lolli,
       height = 24, width = 20, units = "in")


# map of maltreatment in care

p_maltreat_in_care_map <- d |> 
  select(state, year, pct_maltreat_in_care) |>
  filter(state != "Puerto Rico") |> 
  distinct() |>
  ggplot(aes(year, pct_maltreat_in_care)) + 
  geom_line(color = "firebrick") + 
  facet_geo(~ state, 
            grid = "us_state_grid1", 
            scales = "free_y") + 
  map_options + 
  
  #theme(plot.title = element_text(vjust = 1)) +
    
  labs(title = "Prevalence of Maltreatment in Foster Care, 2017-2021", 
       x = "Year",
       y = "Percent of Foster Youth Maltreated", 
       caption = "Source: US Children's Bureau Child Welfare Outcomes Report Data")

ggsave(paste0(fig_path, "pct_maltreated_map.png"), 
       plot = p_maltreat_in_care_map,
       height = 12, width = 20, units = "in")




```



Now for placement instability data

```{r}
# Map of placement instability (percent 3+ placement changes) by year
p_placement_unstable_map <- d |> 
  select(state, year, three_plus_place_all) |> 
  distinct() |> 
  mutate(three_plus_scale = 100 * three_plus_place_all) |> 
  
  filter(state != "Puerto Rico") |> 
  
  ggplot() + 
  geom_line(aes(year, three_plus_scale), 
            color = "firebrick") + 
  facet_geo(~ state, 
            grid = "us_state_grid1", 
            scales = "free_y") + 
  
  labs(title = "Percent Foster Youth with 3+ Placements, 2017-2021", 
       x = "Year", 
       y = "Foster Youth with 3+ Placements", 
       caption = "Source: US Children's Bureau Child Welfare Outcomes Report Data") +
  map_options

ggsave(paste0(fig_path, "place_unstable_map.png"), 
       plot = p_placement_unstable_map, 
       height = 12, width = 20, units = "in")


```


```{r}
# d |> 
#   select(state, year, region, pct_maltreat_in_care, three_plus_place_all) |> 
#   
#   ggplot(aes(pct_maltreat_in_care, three_plus_place_all, 
#              color = as.factor(year))) + 
#   geom_point() + 
#   facet_geo(~ state) +
#   geom_smooth(method = "lm", 
#               se = FALSE)

```


